---
title: 'STOR 455 Homework #9'
subtitle: "25 points - Due 11/28 at 11:59pm"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

__Situation:__ Can we predict income by education level? For this assignment, you will practice building polynomial models through predicting income by education level using data from Elections16 in the Stat2Data dataset. 
 
__Data Source:__ To get the election data, begin with the Election16 data from Stat2Data. The data contains information from all 50 states and the District of Columbia for the 2016 U.S. presidential election. It is similar to Election08 for the 2008 election. 

#### Polynomial Models: ####
1. Calculate the least squares regression line (linear model) that best fits your data using Percent college grads (BA) as the predictor and Income as the response. Interpret (in context) what the slope estimate tells you about income and percent college grads. Include a summary output for your model.

2. Fit a quadratic model using Percent college grads (BA) to predict Income and examine the residuals. Construct a scatterplot of the data with the quadratic fit included. Include a brief discussion of the linear model conditions. 

3. Did the fit improve significantly in question 2 by including the quadratic term? Justify your answer.

# ============================================
# Set-up
# ============================================
```{r}
library(Stat2Data)
library(metRology)
library(tidyr)
library(corrplot)
library(leaps)
library(car)
library(polynom)
```

# ============================================
# Problems
# ============================================
```{r}
data(Election16)
```
# Question 1
1. Calculate the least squares regression line (linear model) that best fits your data using Percent college grads (BA) as the predictor and Income as the response. Interpret (in context) what the slope estimate tells you about income and percent college grads. Include a summary output for your model.
```{r}
modq1 = lm(Income~BA, data = Election16)
summary(modq1)
```
The slope estimate tells us that for every percent increase of college grads in the state, the per capita income in the state is expected to increase by $1467.

# Question 2
2. Fit a quadratic model using Percent college grads (BA) to predict Income and examine the residuals. Construct a scatterplot of the data with the quadratic fit included. Include a brief discussion of the linear model conditions.
```{r}
modq2 = lm(Income~BA + I(BA^2), data = Election16)

plot(Income~BA, data = Election16)

B0 = summary(modq2)$coeff[1,1]
B1 = summary(modq2)$coeff[2,1]
B2 = summary(modq2)$coeff[3,1]

curve(B0 + B1*x + B2*x^2, add = T)

plot(modq2, c(1,2))
```
Linearity seems to be met as our residuals vs fitted plot has a slope close to zero (with some minor curvature in the middle of the data that may be due to the outliers in the middle portion of the plot). Constant variance may be an issue as there seems to be more variability in the residuals as the fitted values increase (evidenced by the fanning pattern). In the normal qq-plot, there is a right skew where there are large deviations from the qqline on the right tail of the qqplot. Normality is not met.

# Question 3
3. Did the fit improve significantly in question 2 by including the quadratic term? Justify your answer.
```{r}
anova(modq1, modq2)
```
No, it did not significantly improve fit. The nested F-test of the two models produced a p-value of 0.9769 - there is not statistically significant evidence to suggest that the coefficient of BA^2 is nonzero. 