---
title: 'STOR 455 Homework #9'
subtitle: "25 points - Due 11/28 at 11:59pm"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

# ============================================
# Set-up
# ============================================
```{r}
library(Stat2Data)
library(metRology)
library(mosaic)
library(tidyr)
library(corrplot)
library(leaps)
library(car)
library(polynom)
library(readr)
library(dplyr)
library(bestglm)
library(MASS)
source("https://raw.githubusercontent.com/JA-McLean/STOR455/master/scripts/ShowSubsets.R")
source("https://raw.githubusercontent.com/JA-McLean/STOR455/master/scripts/anova455.R")
```

# ============================================
# Problems
# ============================================

2. Fit a quadratic model using Percent college grads (BA) to predict Income and examine the residuals. Construct a scatterplot of the data with the quadratic fit included. Include a brief discussion of the linear model conditions. 
```{r}
data(Election16)
plot(Income~BA, data = Election16)
q2 = lm(Income~BA + I(BA^2), data = Election16)

B0 = summary(q2)$coeff[1,1]
B1 = summary(q2)$coeff[2,1]
B2 = summary(q2)$coeff[3,1]

curve(B0 + B1*x + B2*x^2, add = T)

plot(q2)
```

Since the scatterplot already had a generally linear relationship, our quadratic model followed. Our residuals vs fitted values has a plot with a slope close to zero and a qq-plot whose values skew to the right; this means our model does not show normal distribution.